---
output: github_document
---

```{r setup, include = FALSE}
print_yaml <- function(filename) {
  cat("```yaml", readLines(filename), "```", sep = "\n")
}
```

- [Quickstart CI](#quickstart-ci-workflow) - A simple CI workflow to check with the release version of R.
- [Tidyverse CI](#tidyverse-ci-workflow) - A more complex CI workflow
- [Pull Request Commands](#commands-workflow) - Adds `/document` and `/style` commands for pull requests.
- [Render README](#render-readme) - Render README.Rmd when it changes and commit the result
- [Build pkgdown site](#build-pkgdown-site) - Build a [pkgdown] site for an R package and deploy it to [GitHub Pages].
- [Build bookdown site](#build-bookdown-site) - Build a [bookdown] site and deploy it to [netlify].
- [Build blogdown site](#build-blogdown-site) - Build a [blogdown] site and deploy it to [netlify].
- [Managing secrets](#managing-secrets) - How to generate auth tokens and make them available to actions.

## Quickstart CI workflow

This workflow installs latest release R version on macOS
and runs R CMD check via the [rcmdcheck](https://github.com/r-lib/rcmdcheck)
package.

### When can it be used?

1. You have a simple R package
2. There is no OS-specific code
3. You want a quick start with R CI

```{r echo = FALSE, results = "asis"}
print_yaml("check-release.yaml")
```

## Tidyverse CI workflow

This workflow installs the last 5 minor R versions
and runs R CMD check via the [rcmdcheck](https://github.com/r-lib/rcmdcheck)
package on the three major OSs (linux, macOS and Windows). This workflow is
what the tidyverse teams uses on their repositories, but is overkill
for less widely used packages, which are better off using the simpler
quickstart CI workflow.

## When it can be used?

1. You have a complex R package
2. With OS-specific code
3. And you want to ensure compatibility with older R versions

```{r echo = FALSE, results = "asis"}
print_yaml("check-full.yaml")
```

## Commands workflow

This workflow enables the use of 2 R specific commands in pull request issue
comments. `/document` will use [roxygen2](https://roxygen2.r-lib.org/) to
rebuild the documentation for the package and commit the result to the pull
request. `/style` will use [styler](https://styler.r-lib.org/) to restyle your
package.

## When it can they be used?

1. You get frequent pull requests, often with documentation only fixes.
2. You regularly style your code with styler, and require all additions be
   styled as well.

```{r echo = FALSE, results = "asis"}
print_yaml("pr-commands.yaml")
```

## Render README

This example automatically re-builds this README.md from README.Rmd whenever it
or its yaml dependencies change and commits the results to the master branch.

```{r echo = FALSE, results = "asis"}
print_yaml("render-readme.yaml")
```

## Build pkgdown site

This example builds a [pkgdown] site for a repository and pushes the built package
to [GitHub Pages]. **Note** you need to add a `DEPLOY_PAT` secret to your
repository, however this personal access token _only_ needs the `public_repo`
scope.

```{r echo = FALSE, results = "asis"}
print_yaml("pkgdown.yaml")
```

## Build bookdown site

This example builds a [bookdown] site for a repository and then deploys the book via [netlify].
It uses [renv] to ensure the package versions remain consistent across builds.
You will need to run `renv::snapshot()` locally and commit the `renv.lock` file before using this workflow, see [Using renv with Continous Integeration](https://rstudio.github.io/renv/articles/ci.html) for additional information.
**Note** you need to add a `NETLIFY_AUTH_TOKEN` secret to your repository for the netlify deploy (see [Managing secrets](#managing-secrets) section for details).

```{r echo = FALSE, results = "asis"}
print_yaml("bookdown.yaml")
```

## Build blogdown site

This example builds a [blogdown] site for a repository and then deploys the book via [netlify].
It uses [renv] to ensure the package versions remain consistent across builds.
You will need to run `renv::snapshot()` locally and commit the `renv.lock` file before using this workflow, see [Using renv with Continous Integeration](https://rstudio.github.io/renv/articles/ci.html) for additional information.
**Note** you need to add a `NETLIFY_AUTH_TOKEN` secret to your repository for the netlify deploy (see [Managing secrets](#managing-secrets) section for details).

```{r echo = FALSE, results = "asis"}
print_yaml("blogdown.yaml")
```

[GitHub Pages]: https://pages.github.com/
[renv]: https://rstudio.github.io/renv/
[bookdown]: https://bookdown.org
[blogdown]: https://bookdown.org/yihui/blogdown/
[netlify]: https://www.netlify.com/

## Managing secrets

In some cases, your action may need to access an external resource to deploy a result of your action. 
For example, the [bookdown]() and [blogdown]() actions require access to your Netlify account. 
This access is managed using a personal access token, commonly called a PAT.

Netlify has a [process for creating a PAT using their UI](https://docs.netlify.com/cli/get-started/#obtain-a-token-in-the-netlify-ui), which we follow here.

1. In a web browser, open [your Netlify **tokens** page](https://app.netlify.com/user/applications#personal-access-tokens).

2. In another tab in your web browser, open your GitHub repository's **secrets** page. The URL depends on your repository; it will look something like this: `https://github.com/{user}/{repo}/settings/secrets`.

3. At the **tokens** page:

   - Click "New access token". 
   - Provide a description for your benefit, so you will know which token this is, perhaps something like `actions-{repo}`. 
   - Click "Generate token".
   - Copy the token to your clipboard.

4. On your repository's **secrets** page:

   - Click "Add a new secret".
   - In the "Name" field, type `NETLIFY_AUTH_TOKEN` (or the name of the secret that the action expects). 
   - In the "Value" field, paste the token from your clipboard.
   - Click "Add Secret".
   
5. At this point (certainly at some point), you may wish to close your **tokens** page to remove the visibility of your token.   
