---
output: github_document
---

```{r setup, include = FALSE}
print_yaml <- function(filename) {
  cat("```yaml", readLines(filename), "```", sep = "\n")
}
```

## Quickstart CI workflow

This workflow installs latest stable R version on macOS
and runs R CMD check via the [rcmdcheck](https://github.com/r-lib/rcmdcheck)
package.

### When can it be used?

1. You have a simple R package
2. There is no OS-specific code
3. You want a quick start with R CI

```{r echo = FALSE, results = "asis"}
print_yaml("quickstart.yaml")
```

## Tidyverse CI workflow

This workflow installs the last 5 minor R versions
and runs R CMD check via the [rcmdcheck](https://github.com/r-lib/rcmdcheck)
package on the three major OSs (linux, macOS and Windows). This workflow is
what the tidyverse teams uses on their repositories, but is overkill
for less widely used packages, which are better off using the simpler
quickstart CI workflow.

## When it can be used?

1. You have a complex R package
2. With OS-specific code
3. And you want to ensure compatibility with older R versions

```{r echo = FALSE, results = "asis"}
print_yaml("tidyverse.yaml")
```

## Commands workflow

This workflow enables the use of 2 R specific commands in pull request issue
comments. `\document` will use [roxygen2](https://roxygen2.r-lib.org/) to
rebuild the documentation for the package and commit the result to the pull
request. `\style` will use [styler](https://styler.r-lib.org/) to restyle your
package.

## When it can they be used?

1. You get frequent pull requests, often with documentation only fixes.
2. You regularly style your code with styler, and require all additions be
   styled as well.

```{r echo = FALSE, results = "asis"}
print_yaml("commands.yaml")
```

## Re-build README.md every day

This example automatically re-builds the README.md from README.Rmd every
day and opens a PR with the changes (if any). It needs further testing.

```{r echo = FALSE, results = "asis"}
print_yaml("render-readme.yaml")
```
